
# Forecasting bikes demand
Participating in Kaggle competition to forecast bikes sharing demand system, made the following code presenting some supervising methods. here is a link where you can find additional information about the datasets and the competition settings: https://www.kaggle.com/c/bike-sharing-demand


### Loading the data from Kaggle
```{r}
train <- read.csv("..\\train.csv")
test <- read.csv("..\\test.csv")
```

Spliting to train and test datasets
```{r}
train_factor <- train
train_factor$weather <- factor(train$weather)
train_factor$holiday <- factor(train$holiday)
train_factor$workingday <- factor(train$workingday)
train_factor$season <- factor(train$season)

test_factor <- test
test_factor$weather <- factor(test$weather)
test_factor$holiday <- factor(test$holiday)
test_factor$workingday <- factor(test$workingday)
test_factor$season <- factor(test$season)
```

##Data Preparation
Create time column
```{r}
train_factor$time <- substring(train$datetime,12,20)
test_factor$time <- substring(test$datetime,12,20)

train_factor$time <- factor(train_factor$time)
test_factor$time <- factor(test_factor$time)
```


Create day of week column
```{r}
install.packages("lubridate")
library(lubridate)

train_factor$day <- weekdays(as.Date(train_factor$datetime))
train_factor$day <- as.factor(train_factor$day)
test_factor$day <- weekdays(as.Date(test_factor$datetime))
test_factor$day <- as.factor(test_factor$day)
```

Create Sunday variable
```{r}
train_factor$sunday[train_factor$day == "Sunday"] <- "1"                   
train_factor$sunday[train_factor$day != "Sunday"] <- "0"

test_factor$sunday[test_factor$day == "Sunday"] <- "1"
test_factor$sunday[test_factor$day != "Sunday"] <- "0"

#convert to factor
train_factor$sunday <- as.factor(train_factor$sunday)
test_factor$sunday <- as.factor(test_factor$sunday)
```

Create hour variable
```{r}
train_factor$hour<- as.numeric(substr(train_factor$time,1,2))
test_factor$hour<- as.numeric(substr(test_factor$time,1,2))
```
Create daypart column
```{r}
install.packages("car")
library(car)

train_factor$daypart <- "5"
test_factor$daypart <- "5"

#Based on the following graph, daypart bining was made
plot(train_factor$count~train_factor$hour, type = "h")

#23PM - 6AM = 1
train_factor$daypart[(train_factor$hour < 7) | (train_factor$hour > 22)] <- 1
test_factor$daypart[(test_factor$hour < 7) | (test_factor$hour > 22)] <- 1

#7AM - 10AM = 2
train_factor$daypart[(train_factor$hour > 6 ) & (train_factor$hour < 11)] <- 2
test_factor$daypart[(test_factor$hour > 6) & (test_factor$hour < 11)] <- 2

#11AM - 15PM = 3
train_factor$daypart[(train_factor$hour > 10) & (train_factor$hour < 16)] <- 3
test_factor$daypart[(test_factor$hour > 10) & (test_factor$hour < 16)] <- 3

#16PM - 19PM = 4
train_factor$daypart[(train_factor$hour > 15) & (train_factor$hour < 20)] <- 4
test_factor$daypart[(test_factor$hour > 15) & (test_factor$hour < 20)] <- 4

#convert daypart to factor
train_factor$daypart <- as.factor(train_factor$daypart)
test_factor$daypart <- as.factor(test_factor$daypart)

#convert hour back to factor
train_factor$hour <- as.factor(train_factor$hour)
test_factor$hour <- as.factor(test_factor$hour)
```

Making year a new variable
```{r}
train_factor$year <- as.integer(substr(train_factor$datetime,	1,4))
train_factor$year <- as.factor(train_factor$year)

test_factor$year <- as.integer(substr(test_factor$datetime,  1,4))
test_factor$year <- as.factor(test_factor$year)
```
Deleting NVA cols
```{r}
train_factor$datetime <- NULL
train_factor$time <- NULL
```
Finding correlated variables
```{r}
cor(train_factor[c('temp', 'atemp', 'humidity', 'windspeed')])
```

## Data analysis 
Begin with basic libraries and MSE function
```{r}
library(magrittr)
library(dplyr)
MSE <- function(x) x^2 %>% mean
```

OLS
```{r}
ols.1 <- lm(train_factor$count~season + holiday + workingday + weather + temp + atemp + humidity + windspeed + day + sunday + hour + daypart ,data = train_factor)
summary(ols.1)
# R squared is 0.6333

# Train error:
MSE( predict(ols.1)- train_factor$count)
#> 12031.5

# Test error:
MSE( predict(ols.1, newdata = test_factor)- test_factor$count)

```

Random forest
```{r}
library(randomForest)

RandForest <- randomForest(train_factor$count~season + holiday + workingday + weather + atemp + humidity + windspeed + day + sunday + hour + daypart + year, data=train_factor, importance=TRUE, ntree=200, maxnodes = 10, replace = TRUE)

#Plot the variables effect on the model
varImpPlot(RandForest)
```

Predict casual and registered users' demand separatly
```{r}
RandForestCasu.fin <- randomForest(train_factor$casual~season + workingday + weather + temp + humidity + day + hour + daypart + year, data=train_factor, importance=TRUE, ntree=400,mtry = myMtry,  replace = TRUE)
RandForestReg.fin <- randomForest(train_factor$registered~season + workingday + weather + temp + humidity + hour + daypart + year, data=train_factor, importance=TRUE, ntree=400,  mtry = myMtry, replace = TRUE)

RandCasu <- predict (RandForestCasu.fin, test_factor)
RandReg <-predict(RandForestReg.fin, test_factor)

RandForPredict.2 <- round(RandReg + RandCasu, 0)
```
